---
title: "Lab 05 - Wrangling spatial data"
author: "Fiona Wang"
date: "Feb 19"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1
Filter both data frames and only select Alaska. 
```{r dnAK}
dn <- dennys
dn_ak <- dn %>% 
  filter(state == "AK")
nrow(dn_ak)
```

There are 3 Denny's locations in Alaska. 
```{r lqAK}
lq <- laquinta
lq_ak <- lq %>% 
  filter(state == "AK")
nrow(lq_ak)
```

There are 2 La Quinta locations in Alaska. 

### Exercise 2
We want to calculate how many pairs there are.
```{r npair}
npair <- nrow(dn_ak) * nrow(lq_ak)
npair
```

There are 6 pairings between Denny's and La Quinta locations in Alaska. 

### Exercise 3
Let's join the two data frames. 
```{r join, warning=FALSE}
dn_lq_ak <- full_join(dn_ak, lq_ak,
                      by = "state")
dn_lq_ak
```

### Exercise 4
We would like to know how many observations there are in this newly joined data frame, as well as the names of the variables. 
```{r}
nrow(dn_lq_ak)
names(dn_lq_ak)
```
As the two data frames have the same variables, the names have .x or .y to help distinguish. 

### Exercise 5
Haversine
```{r Haversine}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180
  R <- 6371
  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))
  return(round(d, round))
}
```

### Exercise 6
Calculate the distance in each pair.
```{r distance}
dn_lq_ak <- dn_lq_ak %>% 
  mutate(distance = haversine(
    longitude.x, latitude.x, longitude.y, latitude.y, round = 3
  ))
```

Interestingly, I looked up Google map to see if this calculation is correct. However, the distance on google map is always less than what is calculated here. I think it is because we are calculating the sphere, so it's larger.

### Exercise 7
Find the minimum distance between a Denny's and La Quinta for each Denny's location. 
```{r min_distance}
dn_lq_ak_mindist <- dn_lq_ak %>% 
  group_by(address.x) %>% 
  summarize(closest = min(distance))
print(dn_lq_ak_mindist)
```

I realized that these distances are rounded to second.

### Exercise 8
