---
title: "Lab 05 - Wrangling spatial data"
author: "Fiona Wang"
date: "Feb 19"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1
Filter both data frames and only select Alaska. 
```{r dnAK}
dn <- dennys
dn_ak <- dn %>% 
  filter(state == "AK")
nrow(dn_ak)
```

There are 3 Denny's locations in Alaska. 
```{r lqAK}
lq <- laquinta
lq_ak <- lq %>% 
  filter(state == "AK")
nrow(lq_ak)
```

There are 2 La Quinta locations in Alaska. 

### Exercise 2
We want to calculate how many pairs there are.
```{r npair}
npair <- nrow(dn_ak) * nrow(lq_ak)
npair
```

There are 6 pairings between Denny's and La Quinta locations in Alaska. 

### Exercise 3
Let's join the two data frames. 
```{r join, warning=FALSE}
dn_lq_ak <- full_join(dn_ak, lq_ak,
                      by = "state")
dn_lq_ak
```

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...

Add exercise headings as needed.
